Budgeted Reinforcement Learning in Continuous State Space

### 2019-06-11 main.tex
Updates from Overleaf

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage{neurips_2019}

@@

% Paper specific includes and commands
\usepackage{times}
\usepackage{soul}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\urlstyle{same}
\usepackage{xargs}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{import}
\usepackage{svg}
\usepackage{pdfpages}
\usepackage[norelsize, linesnumbered, ruled, lined, boxed, commentsnumbered]{algorithm2e}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{float}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{wrapfig}
\usepackage{enumerate}
\usepackage{minted}

@@

\author{
Nicolas Carrara$^{1,}$\thanks{equal contribution}\and
Edouard Leurent$^{1,3,}$\footnotemark[1]\and
Romain Laroche$^{4}$\and
Tanguy Urvoy$^{2}$\and
Odalric Maillard$^{1}$\and
Olivier Pietquin$^{1,2,}$\thanks{now with Google Research - Brain Team, Paris, France.}\\
\affiliations $^1$SequeL team, INRIA Lille -- Nord Europe, France\\
$^2$Orange Labs, Lannion, France\\
$^3$Renault Group, France\\
$^4$Microsoft Research , Montreal, Canada
\emails
\{nicolas.carrara, edouard.leurent, odalric.maillard, olivier.pietquin\}@inria.fr,
romain.laroche@microsoft.com, tanguy.urvoy@orange.com
}

\begin{document}
\maketitle
\begin{abstract}
    A Budgeted Markov Decision Process (BMDP) is an extension of a Markov Decision Process to critical applications requiring safety constraints. It relies on a notion of risk implemented in the shape of a cost signal constrained  to lie below an -- adjustable -- threshold.
    %A key difference between Constrained Markov Decision Processes (CMDP) and
    %BMDP lies in the ability to change this threshold in real-time.
    %TU: on ne comprend pas cette subtilité à ce stade
    % il vaut mieux développer plus loin.
    % "that -- importantly -- can be modified in real-time."
    So far, BMDPs could only be solved in the case of finite state spaces with known dynamics. This work extends the state-of-the-art to continuous spaces environments and unknown dynamics. We show that the solution to a BMDP is a fixed point of a novel Budgeted Bellman Optimality operator. This observation allows us to introduce natural extensions of Deep Reinforcement Learning algorithms to address large-scale BMDPs. We validate our approach on two simulated applications: spoken dialogue and autonomous driving.
\end{abstract}

@@

Reinforcement Learning (RL) is a general framework for decision-making under uncertainty. It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R_r, \gamma)$ with measurable state space $\cS$, discrete actions $\cA$, unknown rewards $R_r\in\Real^{\cS \times \cA}$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ , where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$. Formally, we seek a policy $\pi\in\cM(A)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$.

@@

Our first contribution is to re-frame the original BMDP formulation in the context of continuous states and infinite discounted horizon. We then propose a novel Budgeted Bellman Optimality Operator and prove the optimal value function to be a fixed point of this operator. Second, we use this operator in \BFTQ, a batch Reinforcement Learning algorithm, for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure. Third, we scale this algorithm to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex programming, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing. Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving. The proofs of our main results are provided \Cref{sec:proofs}.

@@

\begin{theorem}[Contractivity of $\cT$]
\label{thm:contraction}
If $|\cA| \geq 2$, then $\cT$ is not a contraction.
\end{theorem}

@@

The Budgeted Bellman Optimality equation is a fixed-point equation, which motivates the introduction of a fixed-point iteration procedure. We introduce \Cref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs.

Unfortunately, as $\cT$ is not a contraction, we can guarantee neither the convergence of this procedure nor the unicity of its fixed points. Despite those theoretical limitations, we empirically observed the convergence to a fixed point in our experiments (\Cref{sec:experiements}). We conjecture a possible explanation:

@@

When the BMDP is unknown, we need to adapt \Cref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$ collected by interaction with the environment. Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$. We instead use $\hat{\cT}$, a sampling operator, in which this expectation is replaced by:
\begin{equation*}
    \hat{\cT} Q(\os_i, \oa_i, r_i, \os'_i) \eqdef r_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
\end{equation*}
We introduce in \Cref{algo:bftq} the \emph{Budgeted-Fitted-Q} (BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs. Because we work with  continuous state space $\cS$ and budget space $\cB$, we need to employ function-approximation in order to generalise to nearby states and budgets. Precisely, given a parametrized model $Q_\theta$, we seek to minimise a regression loss $\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, r, \os')||_2^2$.
Any model can be used, such as linear models, regression trees, or neural networks.

@@

\section{A Scalable Implementation}
\label{sec:scalable-bftq}
In this section, we introduce an implementation of the BFTQ algorithm designed to operate efficiently and handle large batches of experiences $\mathcal{D}$.

@@

\begin{algorithm}
\DontPrintSemicolon
\KwData{$\os=(s,\beta)$, $Q$}
$Q^+\leftarrow \{Q_c > \min \{Q_c(\os,\oa) \text{ s.t. }\oa\in\argmax_{\oa} Q_r(\os,\oa)\} \}$\tcp*[f]{dominated points}\;
$\cF \leftarrow \text{top frontier of }\texttt{convex\_hull}(Q(\os,\ocA) \setminus Q^+)$\tcp*[f]{candidate mixtures}\;
$\cF_Q \leftarrow \cF\cap Q(\os,\ocA)$\;
\For{points $q = Q(\os,\oa)\in\cF_Q$ in clockwise order}{
\uIf{find two successive points $((q_c^1, q_r^1), (q_c^2, q_r^2))$ of $\cF_Q$ such that $q_c^1 \leq \beta < q_c^2$}{
$p \leftarrow (\beta - q_c^1) / (q_c^2 - q_c^1)$\;
\Return the mixture $(1-p)\delta(\oa-\oa^1) + p\delta(\oa-\oa^2)$\;
}}
\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(\os,\oa))$\tcp*[f]{Budget $\beta$ always respected}}
\caption{Convex hull policy $\pi_\text{hull}(\oa|\os; Q)$}
\label{algo:pi_hull}
\end{algorithm}

@@

%\begin{minipage}[t]{0.5\textwidth}
%\vspace{0pt}
\begin{wrapfigure}{r}{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{source/img/pi.pdf}
    \caption{Representation of $\pi_\text{hull}$. When the budget lies between $Q(\os,\oa_1)$ and $Q(\os,\oa_2)$, two points of the top frontier of the convex hull, then the policy is a mixture of these two points.}
    \label{fig:hull}
\end{wrapfigure}
%\end{minipage}

@@

\subsection{Parallel computing}
\label{subsec:parallel-computing}
In a simulated environment, a first process that can be distributed is the collection of samples in the exploration procedure of \Cref{algo:risk-sensitive-exploration}, as $\pi_\text{greedy}$ stays constant within each mini-batch which avoids the need of synchronisation between workers. Second, the main bottleneck of \BFTQ is the computation of the target $\cT Q$. Indeed, when computing $\pi_\text{hull}$ we must perform at each epoch a Graham-scan of complexity $\cO(|\cA||\tilde{\cB}| \log |\cA\tilde{\cB}|)$ per sample in $\mathcal{D}$ to compute the convex hulls of $Q$ (where $\tilde{\cB}$ is a finite discretisation of $\cB$). The resulting total time-complexity is $\cO(\frac{|\mathcal{D}||\cA||\tilde{\cB}|}{1-\gamma} \log |\cA||\tilde{\cB}|)$. This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA,\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass. By using multiprocessing in the computations of $\pi_\text{hull}$, we enjoy a linear speedup.
The full description of our scalable implementation of \BFTQ is recalled in \Cref{algo:bftq_full} in \Cref{sec:bftq-full}.

@@

\paragraph{Exploration strategies}\label{par:ex-explo} We claimed in \Cref{sec:exploration} that a risk-sensitive exploration was required in the setting of BMDPs. We test this hypotheses by confronting our strategy to a classical risk-neutral strategy. The latter is chosen to be a $\epsilon$-greedy policy slowly transitioning from a random to a greedy policy\footnote{We train this greedy policy using \FTQ.} that aims to maximise $\expectedvalue_{\pi} G_r^\pi$ regardless of $\expectedvalue_{\pi} G_c^\pi$. The quality of the resulting batches $\cD$ is assessed by training a BFTQ policy and comparing the resulting performance.

\paragraph{Budgeted algorithms}\label{par:ex-brl} We compare our  scalable \BFTQ algorithm described in \Cref{sec:scalable-bftq} to an \FTQl baseline. This baseline consists in approximating the BMDP by a finite set of CMDPs problems. We solve each of these CMDP using the standard technique of Lagrangian Relaxation: the cost constraint is converted to a soft penalty weighted by a Lagrangian multiplier $\lambda$ in a surrogate reward function: $\max_{\pi} \expectedvalue_{\pi}[G_r^\pi - \lambda G_c^\pi]$. The resulting MDP can be solved by any RL algorithm, and we chose $\FTQ$ for being closest to $\BFTQ$.
In our experiments, a single training of $\BFTQ$ corresponds to 10 trainings of $\FTQ_l$ policies. Each run was repeated $N_{\text{seeds}}$ times. Parameters of the algorithms can be found in \Cref{sec:algorithms-parameters}

@@

We first bring to light the role of risk-sensitive exploration in the \texttt{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available in \Cref{subsec:exploration-examples}}, and the resulting performance of a budgeted policy trained on each batch. The trajectories (orange) in the risk-neutral batch are concentrated along the risky corridor (red) and ignore the safe corridor (green), which results in bad performances in the low-risk regime. Conversely, trajectories in the risk-sensitive batch (blue) are well distributed among both corridors and the corresponding budgeted policy achieves good performance across the whole spectrum of risk budgets.

@@

\begin{figure}[tp]
    \begin{center}
    \includegraphics[width=0.49\linewidth]{source/img/slot-filling}
    \includegraphics[width=0.49\linewidth]{source/img/highway}
    \caption{Performance comparison of \FTQl and BFTQ on \texttt{slot-filling} (left) and \texttt{highway-env}(right) }
    \label{fig:results}
    \end{center}
\end{figure}

In a second experiment displayed in \Cref{fig:results}, we compare the performance of \FTQl to that of \BFTQ in the dialogue and autonomous driving tasks. For each algorithm, we plot the reward-cost trade-off curve. In both cases, BFTQ performs almost as well as \FTQl despite only requiring a single model. All budgets are well-respected on \texttt{slot-filling}, but on \texttt{highway-env} we can observe an underestimation of $Q_c$, since e.g. $\expectedvalue[G_c|\beta=0] \simeq 0.1 $. This underestimation can be a consequence of two approximations: the use of the sampling operator $\hat{\cT}$ instead of the true environmental operator $\cT$, and the use of the neural network function approximation $Q_\theta$ instead of $Q$. 
% This is probably due to \text{highway-env} being actually infeasible for $\beta=0$, since we noticed that about $8\%$ of bad initialisations occur with the front vehicle driving slower than the minimum velocity of the controlled vehicle, leading to either a front collision or a forced change to the opposite lane, both yielding costs.
Still, \BFTQ provides a better control on the expected cost of the policy, than \FTQl. In addition, \BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance. Examples of policy execution can be found in \Cref{sec:bftq-executions}.

@@

%    \section*{Acknowledgments}
%    %
%    This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
%
%    Thanks to Guillaume Gautier, Fabrice Clerot, Xuedong Shang
\clearpage
\bibliographystyle{named}
\bibliography{budgeted_rl}

>>>>

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
\usepackage[preprint,nonatbib]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
%\usepackage{neurips_2019}

@@

% Paper specific includes and commands
\usepackage[round]{natbib}
\usepackage{xcolor}
\usepackage{times}
\usepackage{soul}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\urlstyle{same}
\usepackage{xargs}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{import}
\usepackage{svg}
\usepackage{pdfpages}
\usepackage[norelsize, linesnumbered, ruled, lined, boxed, commentsnumbered]{algorithm2e}
\usepackage{hyperref}

\definecolor{mylinkcolor}{HTML}{e83f6f}
\definecolor{mycitecolor}{HTML}{2980B9}
\definecolor{myurlcolor}{HTML}{d63230}
% \definecolor{myalgocolor}{HTML}{32936f}
\definecolor{myalgocolor}{HTML}{000000}

\hypersetup{
  linkcolor  = mylinkcolor,
  citecolor  = mycitecolor,
  urlcolor   = myurlcolor,
  colorlinks = true,
}

\usepackage{cleveref}
\usepackage{float}
\usepackage{tabularx}
\usepackage{makecell}
%\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{wrapfig}
\usepackage{enumerate}
%\usepackage{minted}
\usepackage{listings}
\lstset{basicstyle=\ttfamily,
  showstringspaces=false,
  commentstyle=\color{gray},
  keywordstyle=\color{blue}
}

@@

\author{
Nicolas Carrara\thanks{Both authors contributed equally.}\\
SequeL team, INRIA Lille -- Nord Europe\thanks{Univ. Lille, CNRS, Centrale Lille, INRIA UMR 9189 - CRIStAL, Lille, France}\\
\texttt{nicolas.carrara@inria.fr}
\And
Edouard Leurent\footnotemark[1]\\
SequeL team, INRIA Lille -- Nord Europe\footnotemark[2]\\
Renault Group, France\\
\texttt{edouard.leurent@inria.fr}
\And
Romain Laroche\\
Microsoft Research, Montreal, Canada\\
\texttt{romain.laroche@microsoft.com}
\And
Tanguy Urvoy\\
Orange Labs, Lannion, France\\
\texttt{tanguy.urvoy@orange.com}
\And
Odalric-Ambrym Maillard\\
SequeL team, INRIA Lille -- Nord Europe\\
\texttt{odalric.maillard@inria.fr}
\And
Olivier Pietquin\\
Google Research - Brain Team\\
SequeL team, INRIA Lille -- Nord Europe\footnotemark[2]\\
\texttt{pietquin@google.com}
}

\begin{document}
\maketitle
\begin{abstract}
    A Budgeted Markov Decision Process (BMDP) is an extension of a Markov Decision Process to critical applications requiring safety constraints. It relies on a notion of risk implemented in the shape of a cost signal constrained  to lie below an -- adjustable -- threshold.
    So far, BMDPs could only be solved in the case of finite state spaces with known dynamics. This work extends the state-of-the-art to continuous spaces environments and unknown dynamics. We show that the solution to a BMDP is a fixed point of a novel Budgeted Bellman Optimality operator. This observation allows us to introduce natural extensions of Deep Reinforcement Learning algorithms to address large-scale BMDPs. We validate our approach on two simulated applications: spoken dialogue and autonomous driving.
\end{abstract}

@@

Reinforcement Learning (RL) is a general framework for decision-making under uncertainty. It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R_r, \gamma)$ with measurable state space $\cS$, discrete actions $\cA$, unknown rewards $R_r\in\Real^{\cS \times \cA}$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ , where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$. Formally, we seek a policy $\pi\in\cM(\cA)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$.

@@

Our first contribution is to re-frame the original BMDP formulation in the context of continuous states and infinite discounted horizon. We then propose a novel Budgeted Bellman Optimality Operator and prove the optimal value function to be a fixed point of this operator. Second, we use this operator in \BFTQ, a batch Reinforcement Learning algorithm, for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure. Third, we scale this algorithm to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex programming, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing. Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving. The proofs of our main results are provided in \Cref{sec:proofs}.

@@

@@

The Budgeted Bellman Optimality equation is a fixed-point equation, which motivates the introduction of a fixed-point iteration procedure. We introduce \Cref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs. If it were to converge to a unique fixed point, this algorithm would provide a way to compute $Q^*$ and recover the associated optimal budgeted policy $\pi_\text{greedy}(\cdot~; Q^*)$.

\begin{theorem}[Non-contractivity of $\cT$]
\label{thm:contraction}
For any BMDP ($\cS,\cA,P,R_r,R_c,\gamma$) with $|\cA| \geq 2$, $\cT$ is not a contraction. Precisely: $\forall\epsilon>0, \exists Q^1,Q^2\in(\Real^2)^{\ocS\ocA}:\|\cT Q^1-\cT Q^2\|_\infty \geq \frac{1}{\epsilon}\|Q^1-Q^2\|_\infty$.
\end{theorem}

Unfortunately, as $\cT$ is not a contraction, we can guarantee neither the convergence of \Cref{algo:bvi} nor the unicity of its fixed points. Despite those theoretical limitations, we empirically observed the convergence to a fixed point in our experiments (\Cref{sec:experiements}). We conjecture a possible explanation:

@@

When the BMDP is unknown, we need to adapt \Cref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$ collected by interaction with the environment. Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$. We instead use $\hat{\cT}$, a sampling operator, in which this expectation is replaced by:
\begin{equation*}
    \hat{\cT} Q(\os_i, \oa_i, r_i, \os'_i) \eqdef r_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
\end{equation*}
We introduce in \Cref{algo:bftq} the \emph{Budgeted-Fitted-Q} (\BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs. Because we work with  continuous state space $\cS$ and budget space $\cB$, we need to employ function-approximation in order to generalise to nearby states and budgets. Precisely, given a parametrized model $Q_\theta$, we seek to minimise a regression loss $\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, r, \os')||_2^2$.
Any model can be used, such as linear models, regression trees, or neural networks.

@@

\section{A Scalable Implementation}
\label{sec:scalable-bftq}
In this section, we introduce an implementation of the \BFTQ algorithm designed to operate efficiently and handle large batches of experiences $\mathcal{D}$.

@@

\begin{algorithm}
\DontPrintSemicolon
\KwData{$\os=(s,\beta)$, $Q$}
$Q^+\leftarrow \{Q_c > \min \{Q_c(\os,\oa) \text{ s.t. }\oa\in\argmax_{\oa} Q_r(\os,\oa)\} \}$\tcp*[f]{dominated points}\;
$\cF \leftarrow \text{top frontier of }\texttt{convex\_hull}(Q(\os,\ocA) \setminus Q^+)$\tcp*[f]{candidate mixtures}\;
$\cF_Q \leftarrow \cF\cap Q(\os,\ocA)$\;
\For{points $q = Q(\os,\oa)\in\cF_Q$ in clockwise order}{
\uIf{find two successive points $((q_c^1, q_r^1), (q_c^2, q_r^2))$ of $\cF_Q$ such that $q_c^1 \leq \beta < q_c^2$}{
$p \leftarrow (\beta - q_c^1) / (q_c^2 - q_c^1)$\;
\Return the mixture $(1-p)\delta(\oa-\oa^1) + p\delta(\oa-\oa^2)$\;
}}
\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(\os,\oa))$\tcp*[f]{budget $\beta$ always respected}}
\caption{Convex hull policy $\pi_\text{hull}(\oa|\os; Q)$}
\label{algo:pi_hull}
\end{algorithm}

@@

%\begin{minipage}[t]{0.5\textwidth}
%\vspace{0pt}
%\begin{wrapfigure}{r}{0.5\textwidth}
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.7\linewidth]{source/img/pi.pdf}
    \caption{Representation of $\pi_\text{hull}$. When the budget lies between $Q(\os,\oa_1)$ and $Q(\os,\oa_2)$, two points of the top frontier of the convex hull, then the policy is a mixture of these two points.}
    \label{fig:hull}
\end{figure}
%\end{wrapfigure}
%\end{minipage}

@@

\subsection{Parallel computing}
\label{subsec:parallel-computing}
In a simulated environment, a first process that can be distributed is the collection of samples in the exploration procedure of \Cref{algo:risk-sensitive-exploration}, as $\pi_\text{greedy}$ stays constant within each mini-batch which avoids the need of synchronisation between workers. Second, the main bottleneck of \BFTQ is the computation of the target $\cT Q$. Indeed, when computing $\pi_\text{hull}$ we must perform at each epoch a Graham-scan of complexity $\cO(|\cA||\tilde{\cB}| \log |\cA\tilde{\cB}|)$ per sample in $\mathcal{D}$ to compute the convex hulls of $Q$ (where $\tilde{\cB}$ is a finite discretisation of $\cB$). The resulting total time-complexity is $\cO(\frac{|\mathcal{D}||\cA||\tilde{\cB}|}{1-\gamma} \log |\cA||\tilde{\cB}|)$. This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass. By using multiprocessing in the computations of $\pi_\text{hull}$, we enjoy a linear speedup.
The full description of our scalable implementation of \BFTQ is recalled in \Cref{algo:bftq_full} in \Cref{sec:bftq-full}.

@@

\paragraph{Exploration strategies}\label{par:ex-explo} We claimed in \Cref{sec:exploration} that a risk-sensitive exploration was required in the setting of BMDPs. We test this hypotheses by confronting our strategy to a classical risk-neutral strategy. The latter is chosen to be a $\epsilon$-greedy policy slowly transitioning from a random to a greedy policy\footnote{We train this greedy policy using \FTQ.} that aims to maximise $\expectedvalue_{\pi} G_r^\pi$ regardless of $\expectedvalue_{\pi} G_c^\pi$. The quality of the resulting batches $\cD$ is assessed by training a \BFTQ policy and comparing the resulting performance.

\paragraph{Budgeted algorithms}\label{par:ex-brl} We compare our  scalable \BFTQ algorithm described in \Cref{sec:scalable-bftq} to an \FTQl baseline. This baseline consists in approximating the BMDP by a finite set of CMDPs problems. We solve each of these CMDP using the standard technique of Lagrangian Relaxation: the cost constraint is converted to a soft penalty weighted by a Lagrangian multiplier $\lambda$ in a surrogate reward function: $\max_{\pi} \expectedvalue_{\pi}[G_r^\pi - \lambda G_c^\pi]$. The resulting MDP can be solved by any RL algorithm, and we chose \FTQ for being closest to \BFTQ.
In our experiments, a single training of \BFTQ corresponds to 10 trainings of \FTQl policies. Each run was repeated $N_{\text{seeds}}$ times. Parameters of the algorithms can be found in \Cref{sec:algorithms-parameters}

@@

We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available in \Cref{subsec:exploration-examples}}, and the resulting performance of a budgeted policy trained on each batch. The trajectories (orange) in the risk-neutral batch are concentrated along the risky corridor (red) and ignore the safe corridor (green), which results in bad performances in the low-risk regime. Conversely, trajectories in the risk-sensitive batch (blue) are well distributed among both corridors and the corresponding budgeted policy achieves good performance across the whole spectrum of risk budgets.

@@

\begin{figure}[tp]
    \begin{center}
    \includegraphics[width=0.49\linewidth]{source/img/slot-filling}
    \includegraphics[width=0.49\linewidth]{source/img/highway}
    \caption{Performance comparison of \FTQl and \BFTQ on \text{slot-filling} (left) and \text{highway-env}(right) }
    \label{fig:results}
    \end{center}
\end{figure}

In a second experiment displayed in \Cref{fig:results}, we compare the performance of \FTQl to that of \BFTQ in the dialogue and autonomous driving tasks. For each algorithm, we plot the reward-cost trade-off curve. In both cases, \BFTQ performs almost as well as \FTQl despite only requiring a single model. All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can observe an underestimation of $Q_c$, since e.g. $\expectedvalue[G_c|\beta=0] \simeq 0.1 $. This underestimation can be a consequence of two approximations: the use of the sampling operator $\hat{\cT}$ instead of the true environmental operator $\cT$, and the use of the neural network function approximation $Q_\theta$ instead of $Q$. 
% This is probably due to \text{highway-env} being actually infeasible for $\beta=0$, since we noticed that about $8\%$ of bad initialisations occur with the front vehicle driving slower than the minimum velocity of the controlled vehicle, leading to either a front collision or a forced change to the opposite lane, both yielding costs.
Still, \BFTQ provides a better control on the expected cost of the policy, than \FTQl. In addition, \BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance. Examples of policy execution can be found in \Cref{sec:bftq-executions}.

@@

    \section*{Acknowledgments}
    %
    This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR). We thank Guillaume Gautier, Fabrice Clerot, Xuedong Shang for the helpful discussions and valuable insights.
%\clearpage
\bibliographystyle{named}
\bibliography{budgeted_rl}

### 2019-06-11 supplementary.tex
Updates from Overleaf

\begin{center}
\LARGE Appendix
\end{center}
\appendix

@@

Let $(\cS, \cA, P, R_r, R_c)$ be a BMDP.
For any $0 < \epsilon < 1$, we define $Q_\epsilon^1$ and $Q_\epsilon^2$ as:

\begin{align*}
      Q_\epsilon^1(\os,\oa) =
      \begin{cases}
    (0, 1), & \text{if } a = a_0 \\
    (\frac{1}{\epsilon}, 1+\epsilon), & \text{if } a \neq a_0
  \end{cases}\\
  Q_\epsilon^2(\os,\oa) =
      \begin{cases}
    (1, 0), & \text{if } a = a_0 \\
    (1+\frac{1}{\epsilon}, \epsilon), & \text{if } a \neq a_0
  \end{cases}
\end{align*}
Then, $\|Q_1-Q_2\|_\infty = 1$.
 $Q_\epsilon^1$ and $Q_\epsilon^2$ are represented in \Cref{fig:concavity_example}.

@@

But for $\oa=(a,\beta_a)$ with $\beta_a = 1+\epsilon/2$, we have:
\begin{align*}
    \|\cT Q_\epsilon^1(\os, \oa) - \cT Q_\epsilon^2(\os, \oa)\|_\infty &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)} \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^1_\epsilon)}Q^1_\epsilon(\os',\oa') - \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^2_\epsilon)}Q^2_\epsilon(\os',\oa')\right\|_\infty \\
    &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)}(\frac{1}{2\epsilon}, 1+\frac{\epsilon}{2}) - (1+\frac{1}{\epsilon},\epsilon )\right\|_\infty \\
    &= \gamma(1+\frac{1}{2\epsilon})
\end{align*}
Hence, $\|\cT Q_\epsilon^1 - \cT Q_\epsilon^2\|_\infty \geq \gamma(1+\frac{1}{2\epsilon})$

As a conclusion, there does not exist $L>0$ such that:
$$\forall Q_1,Q_2\in(\Real^2)^{\ocS\ocA}, \|\cT Q^1 - \cT Q^2\|_\infty \leq L \|Q^1 - Q^2\|_\infty$$
In other words, $\cT$ is not a contraction for $\|\cdot\|_\infty$.
\end{proof}

@@

\begin{minted}{sh}
# Install highway-env
pip3 install --user git+https://github.com/eleurent/rl-agents
# Change python path to the path of this repository
export PYTHONPATH="code/scaling-up-brl"
# Navigate to budgeted-rl folder
cd code/scaling-up-brl/budgeted-rl/
# Run main script using any config file
# Choose the range of seeds you want to test on
python3 main/egreedy/main-egreedy.py config/slot-filling.json 0 6
python3 main/egreedy/main-egreedy.py config/corridors.json 0 4
python3 main/egreedy/main-egreedy.py config/highway-easy.json 0 10
\end{minted}
\caption{Instructions to reproduce experiments}
    \label{fig:instructions}
\end{figure}

>>>>

\begin{center}
\LARGE Appendices
\end{center}
\appendix

@@

Let $(\cS, \cA, P, R_r, R_c)$ be a BMDP.
For any $\epsilon > 0$, we define $Q_\epsilon^1$ and $Q_\epsilon^2$ as:

\begin{align*}
      Q_\epsilon^1(\os,\oa) =
      \begin{cases}
    (0, 0), & \text{if } a = a_0 \\
    \left(\frac{1}{\gamma}, \epsilon\right), & \text{if } a \neq a_0
  \end{cases}\\
  Q_\epsilon^2(\os,\oa) =
      \begin{cases}
    (0, \epsilon), & \text{if } a = a_0 \\
    \left(\frac{1}{\gamma}, 2\epsilon\right), & \text{if } a \neq a_0
  \end{cases}
\end{align*}
Then, $\|Q_1-Q_2\|_\infty = \epsilon$.
 $Q_\epsilon^1$ and $Q_\epsilon^2$ are represented in \Cref{fig:concavity_example}.

@@

But for $\oa=(a,\beta_a)$ with $\beta_a = \epsilon$, we have:
\begin{align*}
    \|\cT Q_\epsilon^1(\os, \oa) - \cT Q_\epsilon^2(\os, \oa)\|_\infty &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)} \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^1_\epsilon)}Q^1_\epsilon(\os',\oa') - \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^2_\epsilon)}Q^2_\epsilon(\os',\oa')\right\|_\infty \\
    &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)}\left(\frac{1}{\gamma}, \epsilon\right) - (0, \epsilon)\right\|_\infty \\
    &= \gamma\frac{1}{\gamma} = 1
\end{align*}
Hence, 
\begin{align*}
    \|\cT Q_\epsilon^1 - \cT Q_\epsilon^2\|_\infty &\geq 1 = \frac{1}{\epsilon} \|Q_1-Q_2\|_\infty
\end{align*}

In particular, there does not exist $L>0$ such that:
$$\forall Q_1,Q_2\in(\Real^2)^{\ocS\ocA}, \|\cT Q^1 - \cT Q^2\|_\infty \leq L \|Q^1 - Q^2\|_\infty$$
In other words, $\cT$ is not a contraction for $\|\cdot\|_\infty$.
\end{proof}

@@

%\begin{minted}{sh}
\begin{lstlisting}[language=bash,caption={bash version}]
# Install highway-env
pip3 install --user git+https://github.com/eleurent/rl-agents
# Change python path to the path of this repository
export PYTHONPATH="code/scaling-up-brl"
# Navigate to budgeted-rl folder
cd code/scaling-up-brl/budgeted-rl/
# Run main script using any config file
# Choose the range of seeds you want to test on
python3 main/egreedy/main-egreedy.py config/slot-filling.json 0 6
python3 main/egreedy/main-egreedy.py config/corridors.json 0 4
python3 main/egreedy/main-egreedy.py config/highway-easy.json 0 10
\end{lstlisting}
%\end{minted}

\caption{Instructions to reproduce experiments}
    \label{fig:instructions}
\end{figure}
